VM-Centric Snapshot Deduplication\\ for Cloud Data Backup

### 2015-05-07 snapshot_system/mass_arch.tex
re

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to the virtual disk.
Every bit in the bitmap represents a fixed-sized (2MB) segment, indicating whether the segment
has been modified since last backup. 
Segments are further divided into chunks 
%(average 4KB) 
using a content-based chunking algorithm~\cite{frame05}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
For data modification during backup, copy-on-write protection is set so that backup can continue to
copy  a specific version while new changes are recorded.
%copies a frozen version while the modification can still carry on.

@@

Each non-PDS container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
%MSST-CHANGE-BEGIN
For the experiments, we use LZO2 which is a lossless compression libary and does not affect snapshot
data availability. 
%MSST-CHANGE-END
When accessing a particular chunk, its chunk group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
%MSST-CHANGE-BEGIN
A  typical chunk group contains 1000 chunks in our experiment and this setting is effective
for VM snapshot backup we are studying here.
% with an average size of 200-600 chunks.
%MSST-CHANGE-END
%\item
Each non-PDS data container is represented by three files:
%in the DFS
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

@@

%MSST-CHANGE-BEGIN
It should be emphasized that the deleted chunks are marked in the log but do not get deleted until compaction
time.  Only compaction process will read and use this information.  
New snapshots should not refer a chunk marked as deletion
and referenced data will never enter the delete log. This is guaranteed 
by the no-false-negative property of the bloom filter.
%MSST-CHANGE-END
\end{itemize}

>>>>

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to the virtual disk.
Every bit in the bitmap represents a fixed-sized (2MB) segment, indicating whether the segment
has been modified since last backup. 
Segments as super-chunks are further divided into chunks 
%(average 4KB) 
using a content-based chunking algorithm~\cite{frame05}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
For data modification during backup, copy-on-write protection is set so that backup can continue to
copy  a specific version while new changes are recorded.
%copies a frozen version while the modification can still carry on.

@@

Each non-PDS container is further divided into a set of chunk groups for compression. Each chunk compression group is composed of
a set of data chunks and is the basic unit for compression and for data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
%MSST-CHANGE-BEGIN
For the experiments, we use LZO2 which is a lossless compression library and does not affect snapshot
data availability. 
%MSST-CHANGE-END
When accessing a particular chunk, its chunk compression group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
%MSST-CHANGE-BEGIN
A  typical chunk compression group contains 1000 chunks in our experiment and this setting is effective
for the VM snapshot dataset  experienced at Alibaba.
% with an average size of 200-600 chunks.
%MSST-CHANGE-END
%\item
Each non-PDS data container is represented by three files:
%in the DFS
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

@@

%MSST-CHANGE-BEGIN
It should be emphasized that the deleted chunks are marked in the log, but they do not get deleted until compaction
time.  Only compaction process will read and use this information.  
New snapshots should not refer a chunk marked as deletion
and referenced data will never enter the delete log. 
%This is guaranteed by the no-false-negative property of the bloom filter.
%MSST-CHANGE-END
\end{itemize}

### 2015-05-07 snapshot_system/mass_background.tex
re

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  
%MSST-CHANGE-BEGIN
Source-side deduplication is studied at a chunk or file
level~\cite{Vrable2009,Tan2010} for data backup and can be
integrated  with global deduplication among multiple clients~\cite{Fu2014}.
In this context, the client side deduplication is considered as the source-side
and the client-side computing resource  restriction is not a primary concern.
Our work considers the source-side deduplication in a cloud cluster environment before data is transferred
to a backup storage and the computing resource constraint at the cluster side is a major concern.
%MSST-CHANGE-END

>>>>

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  
%MSST-CHANGE-BEGIN
Source-side deduplication is studied at a chunk or file
level~\cite{Vrable2009,Tan2010} for client data backup and can be
integrated  with global deduplication among multiple clients~\cite{Fu2014}.
In this context, the client side is the source-side
and the client-side computing resource  restriction is not a primary concern.
Our work considers the source-side deduplication in a cloud cluster environment before data is transferred
to a backup storage and the computing resource constraint at the cluster side is a major concern.
%MSST-CHANGE-END

### 2015-05-07 snapshot_system/mass_conc.tex
change for msst
re

\section{Discussions and Conclusions}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is the development and analysis of a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
Local similarity search removes 12.05\% of the original data after using a segment-based
 dirty bit method and among them, 3\% is contributed by the increased similarity search.
Notice that our earlier study~\cite{WeiZhangIEEE} reaches 92\% of what global deduplication can do
with a relatively simpler strategy and thus the new design in this paper makes a
VM-centric scheme much more competitive to a traditional VM-oblivious approach.
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  SRB or a VM-oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.

@@

>>>>

\section{Discussions and Conclusions}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is the development and analysis of a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
Local similarity search removes 12.05\% of the original data after using a segment-based
 dirty bit method and among them, 3\% is contributed by the increased similarity search.
Notice that our earlier study~\cite{WeiZhangIEEE} reaches 92\% of what global deduplication can do
with a relatively simpler strategy and thus the new design in this paper makes a
VM-centric scheme much more competitive to a traditional VM-oblivious approach.
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  SRB or a VM-oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.
%MSST-CHANGE-BEGIN
The erasure code has been shown to be effective  to improve reliability on duplicated data 
storage~\cite{Li2011} and such a technique can be incorporated.
Other interesting future work includes
studying the physical layout of backup images~\cite{SmaldoneWH13} and 
further assessing the benefit of PDS separation.

% and comparing with related work~\cite{Vrable2009,Tan2010,Fu2014}. 
%MSST-CHANGE-END

@@

\section*{Acknowledgments}
{
We would like to thank Yujuan Tan and the anonymous referees for their valuable comments,
Hao Jiang, Xiaogang Li, Yue Zeng, Weicai Chen, and Shikun Tian from Alibaba for their kind support and feedback. 
Wei Zhang has received internship support from Alibaba  for VM backup system development.
This work was supported in part by NSF IIS-1118106.
Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and
do not necessarily reflect the views of the National Science Foundation or Alibaba.
}

### 2015-05-07 snapshot_system/mass_eval.tex
re

%{\bf Settings}.
We have performed a trace-driven study based on a production dataset
from Alibaba Aliyun's cloud platform~\cite{Aliyun} 
with about 2500 VMs, running on 100 physical machines. 
Each machine in the production environment is more powerful than our evaluation cluster
and has 16 cores with 48GB memory.
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage 
system while a user may instruct extra snapshots to be saved.
Each VM has about 40GB of storage  data  on average
including OS and user data disk while the size distribution is highly skewed and the maximum to
the average ratio is close to 40.
Each physical machine deals with about 10TB of snapshot data and each 1TB data represents one snapshot version
of 25 VMs.  Thus the total amount of raw snapshot in 100 physical machines
is about 1,000TB.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
%The backup of VM snapshots is required to complete within few  hours every night.
%MSST-CHANGE-BEGIN
Each machine hosts a variety of applications running these operating systems including web services and 
database management. The amount of daily data update is not significant and
the daily snapshot change rate is about 1-3\% on average.
%MSST-CHANGE-END
%All data are divided into 2 MB fix-sized segments and 
The fingerprint for variable-sized chunks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
Note that variable sizing is not required in our scheme and we just follow the setting in the previous work.
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

@@

 Table~\ref{tab:vm-availability} shows 
that VC has a significantly  higher availability than VO.
% as the number of failed machines increases from 3 to 20.
%Our VM-centric approach shows significant improvements
%in VM availability.
With $p=100$ and 3 failures,
VO with 69.5\% availability could lose data in 763 VMs 
while VC with 99.7\% only loses data for 8 VMs out of 2500 VMs.
The key reason is that for most data in VC, only a single VM can be affected by
the loss of a single file block
%Since most blocks contain chunks for a single
%VM, VMs can depend on a smaller number of blocks, 
while in VO, the loss of a
single block tends to affect many VMs.
When the number of physical machines failed reaches 20, the availability of VO reaches 0\%
and VC also drops to 1.13\%. This is because even though availability for PDS data remains good, VC still
loses many of the non-PDS blocks given the replication degree for non-PDS is 3.
When $p=1000$, the percentage of failures is then 2\% and VC delivers a meaningful availability 
with 99.62\%, outperforming VO significantly.

>>>>

%{\bf Settings}.
We have performed a trace-driven study based on a production dataset
from Alibaba Aliyun's cloud platform~\cite{Aliyun} 
with about 2500 VMs, running on 100 physical machines. 
Each machine in the production environment is more powerful than our evaluation cluster
and has 16 cores with 48GB memory.
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage 
system while a user may instruct extra snapshots to be saved.
Each VM has about 40GB of storage  data  on average
including OS and user data disk while the size distribution is highly skewed and the maximum to
the average ratio is close to 40.
Each physical machine deals with about 10TB of snapshot data and each 1TB data represents one snapshot version
of 25 VMs.  Thus the total amount of raw snapshot in 100 physical machines
is about 1,000TB.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
%The backup of VM snapshots is required to complete within few  hours every night.
%MSST-CHANGE-BEGIN
Each machine hosts a variety of applications running these operating systems including web services and 
database management. The amount of daily data update is not significant and
the daily snapshot change rate is about 1-3\% on average.
%MSST-CHANGE-END
%All data are divided into 2 MB fix-sized segments and 
The fingerprint for variable-sized chunks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
Note that variable sizing is not required in our scheme. 
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

@@

 Table~\ref{tab:vm-availability} shows 
that VC has a significantly  higher availability than VO.
% as the number of failed machines increases from 3 to 20.
%Our VM-centric approach shows significant improvements
%in VM availability.
With $p=100$ and 3 failures,
VO with 69.5\% availability could lose data in 763 VMs 
while VC with 99.7\% only loses data for 8 VMs out of 2500 VMs.
The key reason is that for most data in VC, only a single VM can be affected by
the loss of a single file block
%Since most blocks contain chunks for a single
%VM, VMs can depend on a smaller number of blocks, 
while in VO, the loss of a
single block tends to affect many VMs.
The availability of PDS data is very high as seen in  Table~\ref{tab:fsb-availability} 
with more  failures. On the other hand, the increasing of failures  gradually affects the availability of non-PDS data. 
Overall speaking,  VC can  tolerate more failures compared to VO. 
When the number of physical machines failed reaches 20, the availability of VO reaches 0\%
and VC also drops to 1.13\%. This is because even though availability for PDS data remains good, VC still
loses many of the non-PDS blocks given the replication degree for non-PDS is 3.
When $p=1000$, the percentage of failures is then 2\% and VC delivers a meaningful availability 
with 99.62\%, outperforming VO significantly.
